---
output:
  github_document:
    fig_width: 6
    fig_height: 6
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
devtools::load_all()
```

# xrgdal

<!-- badges: start -->
<!-- badges: end -->

The goal of xrgdal is to provide user friendly hooks to GDAL. 

This is a first experiment and what user-tools might look like, with objects keeping a live GDAL pointer open for efficient use (like rgdal did). 

Beware, this is very raw and will crash. The code below is just some pathways
I've engineered to work as a proof of concept. The real work is in
[vapour](https://github.com/hypertidy/vapour), which has a friendly UI
[gdalio](https://github.com/hypertidy/gdalio) for raster data, and there are
helpers for abstract logic in [vaster](https://github.com/hypertidy/vaster) and
[gdalwebsrv](https://github.com/hypertidy/gdalwebsrv) and
[grout](https://github.com/hypertidy/grout) that I'm working into a suite of
tools for development of R software and access to data.


I'd like to see R7 classes that reflect all the underlying functionality of GDAL, like there is with [rasterio](https://rasterio.readthedocs.io/), but we don't need everything I expect, just start with the main things to read and reproject rasters for now. 

## Installation

I actually strongly recommend not doing this. It's too early.  

You can install the development version of xrgdal from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mdsumner/xrgdal")
```

## Example

We have a raster data source (it's an XML config around an online tile server, you can just use your own GeoTIFF etc file/s).

This is an "XYZ" image tile server that simply serves up png or jpeg images from
a tile system, there are 17 `$z`oom levels, 3 bands, and each tile is 256x256.

GDAL has this format for reading from this non-georeferenced raster source,
which is this string `imgsrc`.

```{r example}

template <- 
  '<GDAL_WMS>
    <Service name="AGS">
        <ServerUrl>%s</ServerUrl>
    </Service>
    <DataWindow>
        <UpperLeftX>-20037508.34</UpperLeftX>
        <UpperLeftY>20037508.34</UpperLeftY>
        <LowerRightX>20037508.34</LowerRightX>
        <LowerRightY>-20037508.34</LowerRightY>
        <TileLevel>17</TileLevel>
        <TileCountX>1</TileCountX>
        <TileCountY>1</TileCountY>
        <YOrigin>top</YOrigin>
    </DataWindow>
    <Projection>EPSG:3857</Projection>
    <BlockSizeX>256</BlockSizeX>
    <BlockSizeY>256</BlockSizeY>
    <BandsCount>3</BandsCount>
    <MaxConnections>10</MaxConnections>
    <Cache />
</GDAL_WMS>
'
server <- "http://services.arcgisonline.com/ArcGIS/rest/services/NatGeo_World_Map/MapServer/"
imgsrc <- sprintf(template, server)
```

With xrgdal, we can open this raster and make some queries of it.  These are opaque external pointers atm, so there's only some small set of functions that understand what is available. 

```{r queries}
po <- GDALOpen_cpp(imgsrc, 0)
(dms <- xrgdal_RasterSize(po))  ## it's very big!
xrgdal_GetRasterCount(po)  ## there are 3 bands (RGB)

## get the first band and find out how many overviews (zooms)
band1 <- xrgdal_GetRasterBand(po, 1)
xrgdal_GetOverviewCount(band1)
## and what is the blocksize
xrgdal_GetBlockSize(band1)
```

We know its dimensions and we know how many bands, so let's get some values!

```{r values}
smalldims <- c(360, 180)
v <- xrgdal_RasterIO(po, window = c(0, 0, dms, smalldims), bands = 1:3)
str(v)

```

Thankfully (!!) we didn't read all zillions of pixel values, the `window` argument told us where to start `0, 0` (the topleft corner), where to finish  - `dms[1]` pixels to the right, `dms[2]` pixels to the bottom (the right and bottom edge in this case)  - and how many values to actually obtain `360, 180`. It would take an age to scan through such a big image, but because it has overviews (a.k.a 'zoom levels') GDAL is actually smart enough to choose the closest overview to our requested size. 

These are rgb values, and we can do stuff with that. 

```{r rgb}
a <- array(unlist(v), c(smalldims, 3L))/255
plot(NA, xlim = c(0, smalldims[1]), ylim = c(0, smalldims[2]))
rasterImage(aperm(a, c(2, 1, 3)), 0, 0, smalldims[1], smalldims[2])
```


Now, that looks terrible. So, let's find out what our image device would like and go with that. 

```{r image}
par(mar = rep(0, 4), xpd = NA)
smalldims <- rep(max(dev.size("px")), 2L)
a <- array(unlist(xrgdal_RasterIO(po, window = c(0, 0, dms, smalldims), bands = 1:3)), 
           c(smalldims, 3L))/255
plot(NA, xlim = c(0, smalldims[1]), ylim = c(0, smalldims[2]), xpd = NA, asp = 1)
rasterImage(aperm(a, c(2, 1, 3)), 0, 0, smalldims[1], smalldims[2])


```


## Warping to get an obvious image extent



```{r image-usa}
par(mar = rep(0, 4), xpd = NA)
smalldims <- rep(max(dev.size("px")), 2L)
xlim <- c(-120, -70)
ylim <- c(30, 60)
warp <- xrgdal_Warp(po, extent = c(xlim, ylim), dimension = smalldims, projection = "OGC:CRS84")
a <- array(unlist(xrgdal_RasterIO(warp, window = c(0, 0, smalldims, smalldims), bands = 1:3)), 
           c(smalldims, 3L))/255
plot(NA, xlim = xlim, ylim = ylim, asp = 1/cos(mean(ylim) * pi/180))
rasterImage(aperm(a, c(2, 1, 3)), xlim[1], ylim[1], xlim[2], ylim[2])
maps::map(add = TRUE)

```

etc etc



## Warping to get exactly what we want

TBD

## Using resampling and/or matching the native resolution

TBD

## Churning tiles

TBD 

## Merging disparates sources

TBD


```{r churning-tiles, include = FALSE, eval = FALSE}


info <- vapour_raster_info(src)
info[c("extent", "dimXY", "projection", "tilesXY")]

## now get the tiling index, gives us everthing we need to read a tile for our target, either by the warper or by RasterIO
## the native tiling is tiny so we up it
info$tilesXY
BLOCKSIZE <- info$tilesXY * 8 ## 2, 4, 8, 16, 32
index <- grout::tile_index( grout:::grout(info, BLOCKSIZE))
index

ii <-   seq_len(dim(index)[1L])

po <- xrgdal:::GDALOpen_cpp(src, 0)
system.time({
  for (i in seq_along(ii)) {
    tile <- as.matrix(index[ii[i], ])
    
    vals <- xrgdal:::xrgdal_RasterIO(po, window = tile[, c("offset_x", "offset_y", "ncol", "nrow", "ncol", "nrow"), drop = TRUE])
    #    if (i == 1) print(str(vals))
  }
})

system.time({
  for (i in seq_along(ii)) {
    tile <- as.matrix(index[ii[i], ])
    
    vals <- vapour_read_raster_dbl(src, window = tile[, c("offset_x", "offset_y", "ncol", "nrow", "ncol", "nrow"), drop = TRUE])
#    if (i == 1) print(str(vals))
  }
})


system.time({
  for (i in seq_along(ii)) {
    tile <- as.matrix(index[ii[i], ])
    
    vals <- vapour_warp_raster_dbl(src, 
                                   extent = tile[, c("xmin", "xmax", "ymin", "ymax"), drop = TRUE], 
                                   dimension = tile[, c("ncol", "nrow"), drop = TRUE], projection = info$projection)
 #   if (i == 1) print(str(vals))
  }
})

```
